Measuring Bibliometric Coverage of German Law Journals

### 2025-04-22 docs/paper.qmd
Update paper

---
title: Measuring Bibliometric Coverage of German Law Journals
author:
- name: Christian Boulanger
  orcid: 0000-0001-6928-3246
  email: boulanger@lhlt.mpg.de
  affiliation:
  - name: Max Planck Institute for Legal History and Legal Theory
    city: Frankfurt
    country: Germany
    url: https://www.lhlt.mpg.de/
- name: Daniel Fejzo
  email: fejzo@lhlt.mpg.de
  affiliation:
    - name: Max Planck Institute for Legal History and Legal Theory
      city: Frankfurt
      country: Germany
      url: https://www.lhlt.mpg.de/
- name: Christine Rimmert
  orcid: 0009-0001-2972-2061
  email: Christine.Rimmert@fiz-Karlsruhe.de
  affiliation:
    - name: FIZ Karlsruhe – Leibniz-Institut für Informationsinfrastruktur
      city: Karlsruhe 
      country: Germany
      url: https://www.fiz-karlsruhe.de
abstract: This paper systematically investigates how the German landscape of law journals is represented in databases used for bibliometric analyses, in particular the Web of Science, Scopus, and OpenAlex. ...
lang: en
date: 2025-02-19
licence: "CC BY"
code-fold: true
project:
  execute-dir: project
execute: 
  cache: true
zotero: true
funding: "The authors received no specific fundings for this work."
bibliography: references.bib
csl: ieee-with-url.csl
format:
  pdf: 
    toc: true
    code-annotations: false
  html:
    toc: true
    embed-resources: false
---

@@

%pip install -r requirements.txt
```

@@

[^1]: For this ad-hoc experiment, a dataset has been created on constellate.org, consisting of metadata and fulltext of 45233 documents from the German law journals available on this platform: Archiv für Rechts- und Sozialphilosophie, Archiv des öffentlichen Rechts, Archiv für Rechts- und Sozialphilosophie, Archiv für Rechts- und Wirtschaftsphilosophie, Archiv für die civilistische Praxis, Der Staat, KritV, Kritische Justiz, Rabels Zeitschrift für ausländisches und internationales Privatrecht, Zeitschrift für Rechtspolitik, in the period 1900-2023. These journals are not in any way representative for German legal scholarship, but it is unlikely that this sample would contain significantly fewer hits than any other sample of the same size from a complete corpus.

@@

This research question is subdivided in three sub-questions, which are discussed in separate parts of this paper after shortly surveying previous work. The *first question* asks what exactly we are looking for: what set of publications represents the German law journal landscape? Here, were are not aiming at completeness, but rather at a a sample that is sufficiently large and representative enough to make our observations valid. Based on a previous study that surveyed German law professors to construct a ranking of law journals [@gröls2009], we propose a "gold standard" list of German law journals. Once the sample has been established, the *second question* how much data exists for this sample in our target databases. The *third question* is about data quality: it is not enough to observe the pure quantity of records found for our sample - we also have to investigate how complete and accurate this data is, and what questions can be answered based on it, and which cannot.

@@

This paper is a interdisciplinary collaboration between two domain specialists and a bibliometrician, and thus tries to combine different epistemic interests: on one hand, in those questions that require the data we are looking for, and on the other, the questions that the data poses itself. We document all queries and the code we used to produce the tables and graphs in the following sections in our GitHub repository.[^2]

[^2]: https://github.com/mpilhlt/XXX.

@@

The first task in measuring the coverage of German law journals in bibliographic databases is to determine which of the journals to look for, i.e. to create a list of journals that represent the category "German law journals". In this endeavour, we do not aim for completeness. For the purpose of this study, it will suffice to have a sample which is sufficiently large to be representative of the field. At the same time, the selection must be small enough to be checked and enriched with additional metadata (for example, a persistent identifier such as the ISSN) with a reasonable amount of effort. This will allow us to define a list of journals for which, at a minimum, data has to exist in sufficient quality in order to make statements about knowledge production in German legal scholarship. This does not mean that journals that are not on the list are not important for the discipline. We make no statement on the difference of quality of those journals that make it on the list compared to those which do not. We are not interested in finding or providing a metric similar to the "journal impact factor" of German law journals. We are only interested in the quality of bibliometric data on a representative and relevant list of German law journals.

There are generally two ways in which such a list could be assembled. One approach requires no prior knowledge of the field and promises an "objective" way of generating a list. This *data-based* approach relies on the presence of databases which have sufficiently good metadata categories. It seems not unreasonable to assume that, given a database of journals which has labels for the subject matter "law" and the language attribute "German", one would be able derive a list of German law journals. This would include all journals with German language content, i.e., include journals from Germany, Austria, Switzerland and other legal publications in German. This list could then be filtered further down to journals that cover the Federal Republic of Germany or any other country or combinations of countries. However, as we will discuss below, even if this would work, such an approach will always have at least two problems: a) an additional metric would be needed that would allow to limit this list to the really relevant and representative journals and, even more important, b) the list might be affected by a selection bias in the data from which it was derived. We will therefore argue that we need an alternative approach wich is *based on domain expert knowledge* and which might be more reliable for our purposes.

@@

Instead, the next experiment was to probe the bibliometric databases themselves to see whether, based on the available metadata, we could generate a methodologically controlled list of journals. In theory, this could yield a better selection than what we retrieved from the ZDB, since the bibliometric databases aim to focus on research and are less likely to contain non-scholarly communications. The reason for this selection of these three databases (and the exclusion of others, such as Lens, Semantic Scholar or dimension.ai) is that they are available in a normalized schema, at the Kompetenznetzwerk Bibliometrie (Schmidt et al. 2024).[^4]

@@

[^9]: See https://mjl.clarivate.com/collection-list-downloads for the WoS, and <mark>Scopus-inclusion criteria...</mark>

@@

In other words, for our purposes we need a "gold standard" list against which the results of our queries can be compared. The nature of a gold standard in evaluation research is that we do not have "ground truth" in the sense of empirically verifiable observations, but a yard stick, or "good enough" data (Versi 1992).

@@

The internet offers a number of lists of German law journals. They are curated in the sense as they have been assembled by (usually unnamed) domain experts of varying credentials for pedagogical or other reasons. For the purpose of this article, we scraped the contents of the following lists and converted them into csv data:[^12]

@@

-   "Karlsruher Juristische Bibliographie, Verzeichnis der für Periodika verwendeten Abkürzungen": a list of journals curated by the publisher Beck[^13], with 561 entries;
-   juris: Verzeichnis der ausgewerteten juristischen Periodika (Aufsätze): This is the list of journals ingested by the commercial legal information platform "Juris" for their database on law journal articles. The Bayrische Staatsbibliothek has a PDF from 2002[^14] which we used to extract a list of 383 entries;
-   OpenJur - Juristische Fachzeitschriften: openJur[^15] is a database for german jurisdiction and legislation operated by openJur gGmbH, a non-profit organization under german law. It provides a list of German law journals with 173 entries;
-   Jura Recherche: The internet page https://jura-recherche.de provides a research interface for law students. Its list has 252 entries.

@@

The lists share a relatively small number of abbreviations (43). The number of found co-occurrences could probably be increased somewhat by matching the entries of the list with unique idenitifiers such as the ISSN, which would eliminate incorrect matchings, but probably not much.

@@

Be that as it may, for our purposes, this list has similar problems as the previous ones: as this list hinges on the accidental fact whether an entry exists on a journal or not, and relies on the self-categorization, its represenativeness cannot be guaranteed. And as before, the data itself does not provide any way of ranking the results to give us the most relevant titles.

@@

As the authors point out (<mark>Seite?</mark>), previous attempts at ranking german law journals through empirical means remained scarce. Therefore, the authors decided on a methodological setup derived from similiar studies in economics and business administration. In short, their study consists of different rankings of german law journals generated by surveying the individual assessment of german law scholars. First, 1500 email addresses of german law professors, both current and former, as well as research associates were identified by reviewing the webpages of 45 german universities and academic institutions. Next, these individuals were contacted via email and asked to fill out the survey. 248 responses were received, marking a 16,5% response rate.

The survey asked its participants to name their area of expertise, rank five journals of that area they perceived to be of the highest quality, and additionally, to name and rank five journals that are more general in scope and not bound to a specific subdiscipline, again according to perceived quality. Journals over a certain counting threshold were then assigned scores based on their place in the participants ranking as well as the participants academic status (professors were weighted stronger than research associates) and ‘expert factor’, thus creating two different kinds of rankings: one for journals considered general in nature and several others according to specific subdisciplines.

@@

As stated earlier, one possible objection against using the study by Gröls & Gröls for our gold standard is its age. Specifically, there could be the risk of missing relevant journals established after 2009 or not accounting for the fact that more recent rankings would vastly diverge from one fifteen years ago. In fact, for domain experts it seems an obvious problem that journals as e.g. Rechtswissenschaft (est. 2010) or Recht und Zugang (est. 2020), are necessarily not included in our Gold Standard set and thus are not part of our study.

However, we believe that the utility of using Gröls/Gröls data outweighs those objections for two reasons. First, similar to the situation described by Gröls/Gröls in 2009, there is simply not much alternative research to be found that would be more recent and at least equally suited for a gold standard. In other words, having the risk of incomplete data is more favorable than no data at all. Second, we assume the market for german law journals to be a fairly static one, thus deeming it highly likely that a more recent ranking would more or less include the same publications as fifteen years ago, with the exception of the above mentioned journals.

@@

```{python}
#| label: tbl-gold-standard-coverage
#| tbl-cap: 'Coverage of Gold Standard Journals in the bibliometric databases'

@@

Expressed in a more formal metrics, the *recall* of our gold standard against the trivial search[^21] is 0 for WoS, 0.1 for Scopus and 0.6 for OpenAlex. We cannot really compute a *precision* here, which would require some sort way of determining journal relevance. Our search returned journals having the feature "law" and "German", but with no indication of their quality or impact. In contrast, our Gold Standard is based on their perceived 'relevance' among the population of legal scholars in the survey. In standard bibliometric analysis, the 'impact factor' would be derived by measuring the number of references to articles in that journal (citation count). As we show below, this is not really possible because of the lack of reliable citation data. However, at this point of the analysis, we do not take that information into account.

@@

To find out whether the metadata of the 323 journals in OpenAlex enabled some other form of ranking that would allow to compare the OpenAlex list with our Gold Standard list, we used the following, very simple metrics of each journal in the OpenAlex list: *the number of articles* of indexed, the total *number of citations* of all articles, and the average *number of citations per paper*. We then compared the order of the journals ranked by these metrics with that of the gold standard, with mixed results. <mark>Das folgende braucht eine Tabelle und nochmal eine Überarbeitung</mark>When results show that the pure works count shows a max rank of 299 for a gold standard journal, bad (while the top 5 have ranks 1,2,3,7 and 8 which is good), also both citation based methods provide high ranks for many gold standard journals. Thus, these simple metrics show a stark difference to the ranking of our Gold Standard ranking.

Next, we measured how well the gold standard journals is covered in the bibliometric databases. This question is a somewhat different from the previous as the comparison does not depend on any selection based on metadata in those databases. While the first approach is concerned with the equivalence of common features ("law" , "German"), the second addresses the coverage of these recognized sources within the databases in question. Unsurprisingly, the results confirm our finding that OpenAlex is the only bibliometric database that is worth exploring further. @fig-gold-standard-coverage-visual shows the coverage of our Gold Standard in the Web of Science, Scopus, OpenAlex and CrossRef[^22] that visually makes this clear.

[^22]: CrossRef.org is one of the sources from which OpenAlex draws its data - which can be be seen by the fact that every item from the GS which is contained in CrossRef is also contained in OpenAlex.

```{python}
#| label: fig-gold-standard-coverage-visual
#| fig-cap: 'Coverage of the Gold Standard in bibliometric databases'

@@

For completeness, we also checked how well the Gold Standard aligned with the online lists: Matching again on the journal abbreviations (with all the associated problems) we can see the overlaps of the gold standard with those sources in @fig-inclusion-gold-standard-online-lists:

```{python}
#| label: fig-inclusion-gold-standard-online-lists
#| fig-cap: 'Intersections between the Gold Standard and online lists'

@@

While one would have expected close to 100% inclusion of our gold dataset in the various lists, this is not the case. The results vary from 61% (openjur.de) to 82% (jura-recherche.de)

<mark>-\>DF: Nochmal nachschauen, was genau fehlt: https://gitlab.mpcdf.mpg.de/cboul/kb-bibliometry/-/tree/main/../data/comparisons</mark>

<mark>Was schließen wir daraus?</mark>

@@

As stated above, only one of the three bibliometric databases taken into account seems to cover a sufficient amout of journals considered 'relevant' from experts in the field. For this reason, we concentrate on OpenAlex in the following.

Up to this point, our analysis was restricted to the coverage on journal level, where we found that most of the journals from our gold standard set are covered by OpenAlex. For bibliometric analyses, this coverage on journal level is not sufficient. In addition to the pure existence of the journal in the database, the coverage quality is decicive, which is only given if \* all articles of the journal exist in the database (coverage on article level) \* metadata of articles should be available (coverage on metadata level) in a sufficient quality.

@@

-   **Mohr Siebeck**: JuristenZeitung, Archiv fur die civilistische Praxis, Rabels Zeitschrift fur ausländisches und internationales Recht, Archiv des öffentlichen Rechts
-   **Duncker & Humblot**: Der Staat
-   **Nomos**: Kritische Justiz, Kritische Vierteljahresschrift

@@

In a first step we take a look at item level coverage and simply compare the number of publications in OpenAlex with the number of publications in the publisher data. This looks quite promising: numbers are nearly equal (@fig-compare-publisherdata-openalex-work-count).

@@

Simply having almost equal quantities of articles in either set isn't adequate. It is necessary to ensure that the identical elements exist in both. If this isn't done, some items may be absent in one set and others in another, leading to an equal article count but differing sets. Thus, as a subsequent step, we correlated the articles from OpenAlex against those from publisher data, utilizing the DOI available in both sources. @tbl-compare-publisherdata-openalex-doi shows the number of DOIs which are only found in OpenAlex when categorized by journal with the earliest and the most recent year of missing DOIs. When compared to the overall quantity of articles, these nunbers are neglegible, which means that in terms of the DOI coverage, the data quality in OpenAlex seems good.

@@

-   Duplicate DOIs in OpenAlex, which appeared to correspond to the same publication, but were not present in the publisher data. This was observed in cases where a single publication had multiple DOIs, such as self-archived versions, corrections, or where one DOIs had been assigned to a whole for a whole journal section (i.e. "book reviews") 0rather than the individual components of a the section (i.e., the individual reviews).

@@

```{python}
#| label: fig-oa-zdb-coverage-comparison
#| fig-cap: 'Coverage of Gold Standard journals in OpenAlex compared to actual publication periods (simplified).'

@@

sqldf("""
select 
    journal_name as 'Jounal', 
    pubyear-pubyear_openalex as 'Difference', 
    count(*) as 'Number of Articles'
from eva_all_publisherdata  
where openalex_id is not null
group by  journal_name, pubyear-pubyear_openalex
order by  journal_name, pubyear-pubyear_openalex
""")
```

@@

In sum. the publication year metadata in OpenAlex seems to be usable for bibliometric analyses.

@@

The differences in the number of authors of the publications in OpenAlex vs in the publisher data gives a first impression about the correctness of this data. @fig-cmp-numauthors-oa-publisher shows how many articles with a specific difference in author numbers occur (where a positive difference means more authors in OpenAlex than publisher data while a negative difference means authors missing in OpenAlex). The x-axis corresponds to the difference, the y-axis the number of articles with this difference. So the ideal case would be all the journals having one only value at x=0, showing for all their articles there is an equal number of authors in OpenAlex and the publisher data. As figure 12 shows we do reach this ideal case but for most cases the number of authors is equal, differences appear only in rare cases (low y values for x values \<\> 0).

@@

For author affiliations, the publisher data did not provide affiliation information, making it impossible to compare OpenAlex data with publisher data in this regard. We could still perform statistical analyses on the OpenAlex data itself, focusing on articles from journals in the gold standard set. OpenAlex offers two types of affiliation data: the original raw affiliation string, as it appears on the publication, and an institutional ID assigned through OpenAlex's own disambiguation process. Additionally, OpenAlex provides an institution hierarchy, which could be leveraged for further analysis.

However, the fact that only 4% of the publications have at least one institution_id assigned (6% when excluding publications without authors). And even for these few entries, the affiliation data is most probably incomplete as only one affiliation entry is present, interestingly, most of the affiliations available are from an institution of the Max Planck society. There are a few cases in which a raw affiliation string exists, but no \`institution_id \` could be assigned by the disambiguation procedure.[^23] This situation precludes any bibliometric analysis based on affiliation data, effectively rendering evaluations at the institutional level, including comparisons and rankings of institutions, impossible.

[^23]: Examples for cases with uncleaned data show obvious data errors like e.g. 'gegen Brasilien gewonnen haben?' (W4239815844) or affiliation strings like 'studiert Rechtswissenschaft III Frankfurt am Main' (W4249773843).

@@

# Merge with_ref and total on abk
df = pd.merge(total, with_ref, on='abk', how='right').fillna(0)

@@

::: callout-important
The original plot from the notebook looks different, i.e. has more entries in a slightly different order and showings journals with zero citations
:::

![](https://pad.gwdg.de/uploads/ccb4017b-6cd9-4004-b587-2cfcd9fd5735.png)

>>>>

---
title: Measuring Bibliometric Coverage of German Law Journals
author:
- name: Christian Boulanger
  orcid: 0000-0001-6928-3246
  email: boulanger@lhlt.mpg.de
  affiliation:
  - name: Max Planck Institute for Legal History and Legal Theory
    city: Frankfurt
    country: Germany
    url: https://www.lhlt.mpg.de/
- name: Daniel Fejzo
  email: fejzo@lhlt.mpg.de
  affiliation:
    - name: Max Planck Institute for Legal History and Legal Theory
      city: Frankfurt
      country: Germany
      url: https://www.lhlt.mpg.de/
- name: Christine Rimmert
  orcid: 0009-0001-2972-2061
  email: Christine.Rimmert@fiz-Karlsruhe.de
  affiliation:
    - name: FIZ Karlsruhe – Leibniz-Institut für Informationsinfrastruktur
      city: Karlsruhe 
      country: Germany
      url: https://www.fiz-karlsruhe.de
abstract: "Bibliometric analyses are uncommon in German legal scholarship, often attributed to disciplinary skepticism. However, this paper posits that a more fundamental reason is the lack of adequate data coverage in standard bibliometric databases. We systematically investigate the representation of the German law journal landscape in Web of Science (WoS), Scopus, and OpenAlex. To achieve this, we first establish a gold standard list of 52 representative German law journals based on a 2009 expert survey. We then measure the coverage of these journals within the target databases and assess the quality of the available metadata, focusing particularly on OpenAlex due to its superior initial coverage. Our findings confirm the initial hypothesis: coverage in WoS and Scopus is minimal. While OpenAlex includes a larger proportion of the gold standard journals (32/52) and shows reasonable item-level completeness for a subset checked against publisher data, its metadata quality presents significant limitations. Crucially, affiliation data is sparse, and citation data is almost entirely absent, precluding meaningful citation analysis, impact assessment, or reliable institutional evaluation. We conclude that while coverage is improving, particularly through open platforms like OpenAlex, the current state of bibliometric data, especially the lack of citation information, severely restricts the application of standard bibliometric methods to German legal scholarship. We discuss potential reasons for this gap and suggest focusing on enhancing open data repositories and exploring new methods to generate currently unavailable citation data."
lang: en
date: 2025-02-19
licence: "CC BY"
code-fold: true
project:
  execute-dir: project
execute: 
  cache: true
  echo: false
zotero: true
funding: "The authors received no specific fundings for this work."
bibliography: references.bib
csl: ieee-with-url.csl
format:
  docx: 
    toc: false
    code-annotations: none
  pdf: 
    toc: true
    code-annotations: none
  html:
    toc: true
    embed-resources: false
---

@@

# not necessary if you use `uv`to manage the python version and project dependencies
%pip install -r requirements.txt
```

@@

[^1]: For this ad-hoc experiment, a dataset has been created on the (now defunct) platform constellate.org, consisting of metadata and fulltext of 45233 documents from the German law journals available on this platform: Archiv für Rechts- und Sozialphilosophie, Archiv des öffentlichen Rechts, Archiv für Rechts- und Sozialphilosophie, Archiv für Rechts- und Wirtschaftsphilosophie, Archiv für die civilistische Praxis, Der Staat, KritV, Kritische Justiz, Rabels Zeitschrift für ausländisches und internationales Privatrecht, Zeitschrift für Rechtspolitik, in the period 1900-2023. These journals are not in any way representative for German legal scholarship, but it is unlikely that this sample would contain significantly fewer hits than any other sample of the same size from a complete corpus. Unfortunately, for legal reasons, the data for this experiment cannot be shared.

@@

This research question is subdivided in three sub-questions, which are discussed in separate parts of this paper after shortly surveying previous work. The *first question* asks what exactly we are looking for: what set of publications represents the German law journal landscape? Here, were are not aiming at completeness, but rather at a a sample that is sufficiently large (but not too large) and representative enough to make our observations valid. Based on a previous study that surveyed German law professors to construct a ranking of law journals [@gröls2009], we propose a "gold standard" list of German law journals. Once the sample has been established, the *second question* how much data exists for this sample in our target databases. The *third question* is about data quality: it is not enough to observe the pure quantity of records found for our sample - we also have to investigate how complete and accurate this data is, and what questions can be answered based on it, and which cannot.

@@

This paper is a interdisciplinary collaboration between two domain specialists (Boulanger, Fejzo) and a bibliometrician (Rimmert), and thus tries to combine different epistemic interests: on one hand, in those questions that require the data we are looking for, and on the other, the questions that the data poses itself. We document all queries and the code we used to produce the tables and graphs in the following sections in our GitHub repository.[^2]

[^2]: <https://github.com/mpilhlt/bibliometrics-of-german-law-journals.>

@@

The first task in measuring the coverage of German law journals in bibliographic databases is to determine which of the journals to look for, i.e. to create a list of journals that represent the category “German law journals”. In this endeavour, we do not aim for completeness. For the purpose of this study, it will suffice to have a sample which is sufficiently large to be representative of the field. At the same time, the selection must be small enough, so it is possible to check and enrich the dataset with additional metadata (for example, a persistent identifier such as the ISSN) with a reasonable amount of effort. This will allow us to define a list of journals for which, at a minimum, data has to exist in sufficient quality in order to make statements about knowledge production in German legal scholarship. This does not mean that journals that are not on the list are not important for the discipline. In fact, as it will become clear, our selection methodology forces us to omit any journal founded after 2009. We also make no statement on the difference of quality of those journals that make it on the list compared to those which do not. We are not interested in finding or providing a metric similar to the “journal impact factor” of German law journals. We are only interested in the quality of bibliometric data on a representative and relevant list of those journals.

There are generally two ways in which such a list could be assembled. One approach requires no prior knowledge of the field and promises an "objective" way of generating a list. This *data-based* approach relies on the presence of databases which have sufficiently good metadata categories. It seems not unreasonable to assume that, given a database of journals which has labels for the subject matter "law" and the language attribute "German", one would be able derive a list of German law journals. This would include all journals with German language content, i.e., include journals from Germany, Austria, Switzerland and other legal publications in German. This list could then be filtered further down to journals that cover the Federal Republic of Germany or any other country or combinations of countries. However, as we will discuss below, even if this would work, such an approach will always have at least two problems: a) an additional metric would be needed that would allow to limit this list to the really relevant and representative journals and, even more important, b) the list might be affected by a selection bias in the data from which it was derived. We will therefore argue that we need an alternative approach which is *based on domain expert knowledge* and which might be more reliable for our purposes.

@@

Instead, the next experiment was to probe the bibliometric databases themselves to see whether, based on the available metadata, we could generate a methodologically controlled list of journals. In theory, this could yield a better selection than what we retrieved from the ZDB, since the bibliometric databases aim to focus on research and are less likely to contain non-scholarly communications. The reason for this selection of these three databases (and the exclusion of others, such as Lens, Semantic Scholar or dimension.ai) is that they are available in a normalized schema, at the Kompetenznetzwerk Bibliometrie [@schmidtEtAl2024].[^4]

@@

[^9]: See https://mjl.clarivate.com/collection-list-downloads for the WoS.

@@

In other words, for our purposes we need a "gold standard" list against which the results of our queries can be compared. The nature of a gold standard in evaluation research is that we do not have "ground truth" in the sense of empirically verifiable observations, but a yard stick, or "good enough" data [@versi1992].

@@

The internet offers a number of lists of German law journals. They are curated in the sense as they have been assembled by (usually unnamed) domain experts of varying credentials for pedagogical or other reasons. For the purpose of this article, we scraped the contents of the following lists and converted them into CSV data:[^12]

@@

-   "Karlsruher Juristische Bibliographie, Verzeichnis der für Periodika verwendeten Abkürzungen": a list of journals curated by the publisher Beck[^13], with 561 entries;
-   juris: Verzeichnis der ausgewerteten juristischen Periodika (Aufsätze): This is the list of journals ingested by the commercial legal information platform "Juris" for their database on law journal articles. The Bayrische Staatsbibliothek has a PDF from 2002[^14] which we used to extract a list of 383 entries;
-   OpenJur - Juristische Fachzeitschriften: openJur[^15] is a database for german jurisdiction and legislation operated by openJur gGmbH, a non-profit organization under german law. It provides a list of German law journals with 173 entries;
-   Jura Recherche: The internet page <https://jura-recherche.de> provides a research interface for law students. Its list has 252 entries.

@@

The lists share a relatively small number of abbreviations (43). The number of found co-occurrences could probably be increased somewhat by matching the entries of the list with unique identifiers such as the ISSN, which would eliminate incorrect matching, but probably not much.

@@

Be that as it may, for our purposes, this list has similar problems as the previous ones: as this list hinges on the accidental fact whether an entry exists on a journal or not, and relies on the self-categorization, its representativeness cannot be guaranteed. And as before, the data itself does not provide any way of ranking the results to give us the most relevant titles.

@@

As the authors point out, previous attempts at ranking German law journals through empirical means remained scarce (497). Therefore, the authors decided on a methodological setup derived from similar studies in economics and business administration (489). In short, their study consists of different rankings of German law journals generated by surveying the individual assessment of German law scholars. First, 1500 email addresses of German law professors, both current and former, as well as research associates were identified by reviewing the web pages of 45 German universities and academic institutions. Next, these individuals were contacted via email and asked to fill out the survey. 248 responses were received, marking a 16,5% response rate (490).

The survey asked its participants to name their area of expertise, rank five journals of that area they perceived to be of the highest quality, and additionally, to name and rank five journals that are more general in scope and not bound to a specific subdiscipline, again according to perceived quality (490). Journals over a certain counting threshold were then assigned scores based on their place in the participant's ranking as well as the participant's academic status (professors were weighted stronger than research associates) and an ‘expert factor’ determined by the individuals amount of publications over the last three years, thus creating two different kinds of rankings: one for journals considered general in nature and several others according to specific subdisciplines (491-492).

@@

As stated earlier, one possible objection against using the study by Gröls & Gröls for our gold standard is its age. Specifically, there could be the risk of missing relevant journals established after 2009 or not accounting for the fact that more recent rankings would vastly diverge from one fifteen years ago. In fact, for domain experts it seems an obvious problem that journals as e.g. Rechtswissenschaft (est. 2010) or *Recht und Zugang* (est. 2020), are necessarily not included in our gold standard set and thus are not part of our study.

However, we believe that the utility of using Gröls/Gröls data outweighs those objections for two reasons. First, similar to the situation described by Gröls/Gröls in 2009, there is simply not much alternative research to be found that would be more recent and at least equally suited for a gold standard. In other words, having the risk of incomplete data is more favorable than no data at all. Second, we assume the market for German law journals to be a fairly static one, thus deeming it highly likely that a more recent ranking would more or less include the same publications as fifteen years ago, with the exception of the above mentioned journals.

@@

```{python}
#| label: tbl-gold-standard-coverage
#| tbl-cap: 'Coverage of gold standard Journals in the bibliometric databases'

@@

Expressed in a more formal metrics, the *recall* of our gold standard against the trivial search[^21] is 0 for WoS, 0.1 for Scopus and 0.6 for OpenAlex. We cannot really compute a *precision* here, which would require some sort way of determining journal relevance. Our search returned journals having the feature "law" and "German", but with no indication of their quality or impact. In contrast, our gold standard is based on their perceived 'relevance' among the population of legal scholars in the survey. In standard bibliometric analysis, the 'impact factor' would be derived by measuring the number of references to articles in that journal (citation count). As we show below, this is not really possible because of the lack of reliable citation data. However, at this point of the analysis, we do not take that information into account.

@@

To find out whether the metadata of the 323 journals in OpenAlex enabled some other form of ranking that would allow to compare the OpenAlex list with our gold standard list, we used the following, very simple metrics of each journal in the OpenAlex list: *the number of articles* of indexed, the total *number of citations* of all articles, and the average *number of citations per paper*. We then compared the order of the journals ranked by these metrics with that of the gold standard, with mixed results. On one hand, the top 5 on our gold standard rank among the top ten in the number of articles, however, others are placed at the bottom. Some citation based methods provide high ranks for many, but not all gold standard journals. Thus, these simple metrics show a stark difference between conceivable ranking methods in OpenAlex and our gold standard ranking.[^22]

[^22]: <https://github.com/mpilhlt/bibliometrics-of-german-law-journals/blob/main/notebooks/evaluations_rewi_create%20basic_tables.ipynb>

Next, we measured how well the gold standard journals is covered in the bibliometric databases. This question is a somewhat different from the previous as the comparison does not depend on any selection based on metadata in those databases. While the first approach is concerned with the equivalence of common features ("law" , "German"), the second addresses the coverage of these recognized sources within the databases in question. Unsurprisingly, the results confirm our finding that OpenAlex is the only bibliometric database that is worth exploring further. @fig-gold-standard-coverage-visual shows the coverage of our gold standard in the Web of Science, Scopus, OpenAlex and CrossRef[^23] that visually makes this clear.

[^23]: CrossRef.org is one of the sources from which OpenAlex draws its data - which can be be seen by the fact that every item from the GS which is contained in CrossRef is also contained in OpenAlex.

```{python}
#| label: fig-gold-standard-coverage-visual
#| fig-cap: 'Coverage of the gold standard in bibliometric databases'

@@

For completeness, we also checked how well the gold standard aligned with the online lists: Matching again on the journal abbreviations (with all the associated problems) we can see the overlaps of the gold standard with those sources in @fig-inclusion-gold-standard-online-lists:

```{python}
#| label: fig-inclusion-gold-standard-online-lists
#| fig-cap: 'Intersections between the gold standard and online lists'

@@

While one would have expected close to 100% inclusion of our gold dataset in the various lists, this is not the case. The results vary from 61% (openjur.de) to 82% (jura-recherche.de).

@@

As stated above, only one of the three bibliometric databases taken into account seems to cover a sufficient amount of journals considered 'relevant' from experts in the field. For this reason, we concentrate on OpenAlex in the following.

Up to this point, our analysis was restricted to the coverage on journal level, where we found that most of the journals from our gold standard set are covered by OpenAlex. For bibliometric analyses, this coverage on journal level is not sufficient. In addition to the pure existence of the journal in the database, the coverage quality is decisive, which is only given if a) all articles of the journal exist in the database (coverage on article level), and b) metadata of articles should be available (coverage on metadata level) in a sufficient quality.

@@

-   **Mohr Siebeck**: JuristenZeitung; Archiv fur die civilistische Praxis; Rabels Zeitschrift für ausländisches und internationales Recht; Archiv des öffentlichen Rechts
-   **Duncker & Humblot**: Der Staat
-   **Nomos**: Kritische Justiz; Kritische Vierteljahresschrift

@@

In a first step we take a look at item level coverage and simply compared the number of publications in OpenAlex with the number of publications in the publisher data. This looks quite promising: numbers are nearly equal (@fig-compare-publisherdata-openalex-work-count).

@@

Simply having almost equal quantities of articles in either set isn't adequate. It is necessary to ensure that the identical elements exist in both. If this isn't done, some items may be absent in one set and others in another, leading to an equal article count but differing sets. Thus, as a subsequent step, we correlated the articles from OpenAlex against those from publisher data, utilizing the DOI available in both sources. @tbl-compare-publisherdata-openalex-doi shows the number of DOIs which are only found in OpenAlex when categorized by journal with the earliest and the most recent year of missing DOIs. When compared to the overall quantity of articles, these numbers are negligible, which means that in terms of the DOI coverage, the data quality in OpenAlex seems good.

@@

-   Duplicate DOIs in OpenAlex, which appeared to correspond to the same publication, but were not present in the publisher data. This was observed in cases where a single publication had multiple DOIs, such as self-archived versions, corrections, or where one DOIs had been assigned to a whole for a whole journal section (i.e. "book reviews") rather than the individual components of a the section (i.e., the individual reviews).

@@

```{python}
#| label: fig-oa-zdb-coverage-comparison
#| fig-cap: 'Coverage of gold standard journals in OpenAlex compared to actual publication periods (simplified).'

@@

sqldf("""
select 
    journal_name as 'Journal', 
    cast(pubyear-pubyear_openalex AS INTEGER) as 'Difference', 
    count(*) as 'Number of Articles'
from eva_all_publisherdata  
where openalex_id is not null
group by  journal_name, pubyear-pubyear_openalex
order by  journal_name, pubyear-pubyear_openalex
""")
```

@@

In sum, the publication year metadata in OpenAlex seems to be usable for bibliometric analyses.

@@

The differences in the number of authors of the publications in OpenAlex vs in the publisher data gives a first impression about the correctness of this data. @fig-cmp-numauthors-oa-publisher shows how many articles with a specific difference in author numbers occur (where a positive difference means more authors in OpenAlex than publisher data while a negative difference means authors missing in OpenAlex). The x-axis corresponds to the difference, the y-axis the number of articles with this difference. So the ideal case would be all the journals having one only value at x=0, showing for all their articles there is an equal number of authors in OpenAlex and the publisher data. As @fig-cmp-numauthors-oa-publisher shows, we do reach this ideal case but for most cases the number of authors is equal, differences appear only in rare cases (low y values for x values \<\> 0).

@@

For author affiliations, the publisher data did not provide affiliation information, making it impossible to compare OpenAlex data with publisher data in this regard. Thus we limited our analysis to statistical analyses on the OpenAlex data itself, focusing on articles from journals in the gold standard set.

OpenAlex offers two types of affiliation data: the original raw affiliation string, as it appears on the publication, and an institutional ID assigned through OpenAlex's own disambiguation process. Additionally, OpenAlex provides an institution hierarchy, which could be leveraged for further analysis. However, the fact that only 4% of the publications have at least one \`institution_id\` assigned (6% when excluding publications without authors). And even for these few entries, the affiliation data is most probably incomplete as only one affiliation entry is present. Interestingly, most of the affiliations available are from an institution of the Max Planck society. There are a few cases in which a raw affiliation string exists, but no \`institution_id\` could be assigned by the disambiguation procedure.[^24] This situation precludes any bibliometric analysis based on affiliation data, effectively rendering evaluations at the institutional level, including comparisons and rankings of institutions, impossible.

[^24]: Examples for cases with uncleaned data show obvious data errors like e.g. 'gegen Brasilien gewonnen haben?' (W4239815844) or affiliation strings like 'studiert Rechtswissenschaft III Frankfurt am Main' (W4249773843).

@@

# Merge with_ref and total on abk
df = pd.merge(total, with_ref, on='abk', how='left').fillna(0)
df = df.sort_values(by='abk', key=lambda x: x.str.lower())

@@

We also had a look at the reference situation with respect to the publication years to see if more actual publications do have a better coverage of references but this is not the case. @tbl-oa-references-statistics shows that 5,097 distinct articles appear as a reference in at least one article of our gold standard set, of which 41.55% are in the gold standard set of journals; 43.24% of them are publications in journals we found by trivial search (German, law) in OpenAlex. However, it must be kept in mind that these figures do not mean much because the reference data is simply to scarce for any kind of robust interpretation.

```{python}
#| label: tbl-oa-references-statistics
#| tbl-cap: 'Statistics on OpenAlex reference data in the gold standard'
